\chapter{Reflections} \label{cha:reflections}
The first part of this thesis presented the background and five examples for the design of visualization applications that utilize a combination of scientific visualization, information visualization, and visual analytics techniques in the context of the categories of visualization applications, \emph{Exploration}, \emph{Analysis}, and \emph{Communication}.

This final chapter provides a summary of this first part of the thesis, includes personal reflections, and considers possible venues for future work.  The future work includes interesting research for myself, but also contains humble suggestions for the visualization community towards steps that help to continue the accelerating growth of the field.  In-depth discussions for concrete potential future work in each of domain discussed in the previous chapter can be found in the publications attached in the second part of this thesis.

This thesis follows the line of argumentation that emphasizes the importance of considering the intended target audience when designing and creating visualization applications.  This builds on previous research highlighting the difference between applications designed for \emph{explorational} and \emph{presentational} use cases by van Wijk~\cite{van2005value} and the distinction into three (different) categories by Keim \etal\ ~\cite{keim2006challenges} and Schulz \etal ~\cite{schulz2013design}.  The papers included in this thesis illustrate that the explorational use cases presented by van Wijk can be further subdivided into initial exploratory hypothesis generation and repeated analyses.  This leads to the three categories of visualization applications, \emph{Exploration}, \emph{Analysis}, and \emph{Communication} described in Chapter~\ref{cha:intro}. Table~\ref{tab:definitions} shows a comparison of the definitions provided by Keim et al.~\cite{keim2006challenges}, Schulz et al.~\cite{schulz2013design}, Brehmer et al.~\cite{brehmer2013typology}, van Wijk, where they overlap and how they differ.  A common element between all classifications is the distinction between visualization applications that are designed for exploration and applications designed for presentation/communication.  However, I would argue for the need for a more holistic approach for application design.

An essential component of these classifications are the task-based taxonomoies and typologies presented by, among others, Shneiderman, Brehmner \etal , Schulz \etal , and others, some of which take the goal of the visualization task and thus the application directly into account.  However, are multiple deficiencies with this approach.  In many taxonomies, while the differentiation between individual tasks is not always clear, the border between different user groups is much less pronounced.  At most, such as in Miksch \etal Data-Users-Tasks Design Triangle~\cite{miksch2014matter}, the users take second place and and are differenciated by application domain, physical environment, social factors, and technical specifics.  In my opinion, the existing successful taxonomies will have to be adapted to take into account many more variables about the user groups that are impacted by visualization application and task design.  This consideration would ask for a higher granularity of \emph{Communication}-based visualization applications, which Brehmner \etal argues for.

Another aspect of the holistic approach to application design is signified by the quote by Kosara and Mackinlay ``Visualization researchers often tacitly assume that the tools used for analysis are usable for presentation just as well as for their original purpose. We believe that to be a very limiting assumption, however.''~\cite{kosara2013storytelling}.  I would instead argue that this is a very freeing assumption instead, due to the ongoing convergence of visualization applications aimed at \emph{Exploration} and \emph{Communication}.  While there has always been exchanges between visualization applications designed for exploration and analysis, in the past the influence of these two categories on the communication aspect of visualization was less pronounced.  The visualization application presented in Section~\ref{contributions:astro:openspace} shows the potential of utilizing guided exploration for the public dissemination of complex scientific topics and how simplified exploratory methods can be used for the education of the general public while providing highly complex tools for domain scientists in the same framework.  This combination allows scientists to rapidly disseminate their research finding without the active process of converting data into a new format for a different real-time application in order to use their findings for a guided, interactive presentation for the general public.  This convergence of use cases is currently ongoing and induced by the availability of high performance GPUs on commodity hardware that enable the use of real, large scale and complex scientific data in both exploratory scenarios, as well as for communication.  As an additional benefit, the trend on focussing on the accessibility of visualization methods for the general public was present since the beginning of the visualization field, but has been increasing in the past years.  The availability of visualization tools is especially beneficial for an interested general public.  Providing access to sophisticated tools for the general public can provide a huge benefit to citizen science projects which increases the combined human knowledge and scientific engangement simultaneously.  Naturally, this does not argue for the stagnation of fundamental visualization research but rather argues for an increase in the number of visualization researchers (another trend which can be observed at conferences).

Undoubtedly, the science of visualization and its tools have matured a lot in the past 25 years and now has a large influence on the scientific method in almost every field of research.  High-impact scientific publications without accurate accompanying visualizations of their findings are unthinkable today.  Furthermore, in many cases, no visualization experts are necessary to create these final visualizations as the access to sophisticated visualization tools has increased dramatically.  The increasing ease-of-use of existing visualization techniques has thus led to a proliferation of knowledge about visualization techniques.  While this, in my opinion, is a worthwhile trajectory of visualization research as it enables access to these tools to an increasing number of people, it also critically important to increase the users' visualization literacy alongside the tools' ease-of-use.

In my opinion, one of the most interesting and important developments that are currently ongoing is the intersection of visualization and machine learning methods.  One of the most basic concepts of visualization is to connect the best abilities of the human mind with the best abilities of computational resources.  Combining the natural intelligence of a human expert with the artificial intelligence of deep learning methods can be hugely beneficial for the field.  While, there are interesting debates on the viability of artificial intelligence to replace human cognition, these are better left outside the scope of this work.  But until/if that happens, visualization techniques can certainly benefit from the rapid development in deep learning and vice versa.  In the reference frame of visualization as the field of using the human intelligence where it performs best and computational resources where they are most efficient, the introduction of artificial intelligence merely shifts that boundary, rather than removing it completely.  This is illustrated by the quote from Stefan Lindholm: ``One trend is the ever increasing number of arrows used whenever a visualization pipeline is illustrated. The narrow concept of a linear pipeline with a few fixed stages is going extinct''~\cite{lindholm14medical}.  A suitable addition to this quote is the realization that it is important to remember that in the visualization pipeline, all arrows start and end at the human-in-the-loop.

\vfill \hfill \emph{Alexander Bock}